---
title: "Mobility PLots"
author: "Cheyenne, Ziyan"
date: "4/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in the spline slopes

```{r message=FALSE, warning=FALSE, cache=TRUE}
source("step2_data_wrangle.r")
# SHIFT the DATE for mobility as well: mobility a week ago may impact the infections number now
case_mobility$DATE <- case_mobility$DATE - 7 ## WARNING: only run this once
```

# Mobility

Plot1.1: cumulative mobility per color (plot the 3 curves on the same plot).

Aggregate counties by teaching method

```{r}
cum_mobility <- case_mobility%>%
  left_join(wide_teaching_enroll,by=c("COUNTY"="county"))%>%
  left_join(major_reopening,by="COUNTY")%>%
  group_by(DATE,major_teaching)%>%
  summarise(daily_tot_teaching_full_work= sum(full_work_prop_7d*POPULATION),
            tot_teaching_pop = sum(POPULATION), .groups = "drop")%>%
  group_by(major_teaching)%>%
  arrange(DATE)%>%
  mutate(cum_full_work = cumsum(daily_tot_teaching_full_work),cum_full_work_prop = cumsum(daily_tot_teaching_full_work)/tot_teaching_pop)%>%
  mutate(shiftedDate = DATE+28) ## Shifted Dates!!!!!!!!!!!!

cum_mobility$major_teaching <- factor(cum_mobility$major_teaching,levels = c("On Premises","Hybrid","Online Only"))
```


```{r}
options(scipen=1000)

ggplot(cum_mobility, aes(x = shiftedDate, color = major_teaching)) +
  geom_line(aes(y = cum_full_work), alpha = 1,size=1)+ 
  geom_rect(data = cum_mobility[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "orange") + 
  theme_bw() + 
  labs(x = "Date", y = "Cumulative Full-time Workers/Students", 
       color = "Majority Teaching Method", title="Cumulative Full-time Mobility (6hrs+ Away Home)\n4 Weeks Forward",
       subtitle = "Yellow area represents the fall semester (08/18 - 12/15)") + 
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),legend.position = "bottom")+theme(legend.text = element_text(size=12),legend.title = element_text(size=12),axis.text = element_text(size=13),title=element_text(size=13))

#ggsave("cum_mobility_time_series.png",width = 7, height = 5)
```

Plot1.2: average mobility per color (plot the 3 curves on the same plot).

```{r}
date.intercept=as.Date("2020-11-24")

date_mobility$major_teaching <- factor(date_mobility$major_teaching,levels = c("On Premises","Hybrid","Online Only"))

date_mobility$shiftedDate <- date_mobility$DATE + 28

# FULL TIME WORK, 6 hours +
date_mobility%>%
  ggplot(aes(x=shiftedDate,y=full_work_prop))+
  geom_rect(data=date_mobility[1,],
            aes(xmin=as.Date("2020/08/26"), xmax=as.Date("2020/12/12"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "orange") + 
  geom_line(aes(color = major_teaching),size = 1, alpha = .8)+
  geom_vline(xintercept = date.intercept, linetype = "dashed") + 
  annotate("text",x = date.intercept,y = 0.08,
           label = date.intercept,
           hjust = 1.1,size=5) + 
  theme_bw() + 
  labs(x = "Date", y = "% Full-time Workers/Students",subtitle = "Average over 7 days; Yellow area represents Fall Semester",
       color = "Majority Teaching Method",title = "Average %Cell Phones Away Home for 6hr+\n4 Weeks Forward") + ylim(0.02,0.10)+
  theme(legend.position = "bottom")+theme(legend.title = element_text(size=13),legend.text = element_text(size=13),axis.title = element_text(size=15),axis.text = element_text(size=15),legend.background = element_rect(fill = alpha("orange",0.0)),legend.key.size = unit(1.4,"lines"),title = element_text(size=12.9))+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) 

```

\newpage

Then a series of 3 plots, each plot showing the first diff of log death and 

cum mobility for each color. If the 2 curves are parallel, then mobility explains death entirely and school has no effect.

Aggregate counties by teaching method (cases_clean)

Plot2.1: On Premises

```{r}
cases_slope_teach_agg%>%
  filter(major_teaching=="On Premises")%>%
  inner_join(cum_mobility%>%filter(major_teaching=="On Premises"),by=c("DATE"="shiftedDate"))%>%
  ggplot(aes(x = DATE))+
  geom_line(aes(y = B),size = 1,color="red") +
  geom_rect(data = cases_slope_teach_agg[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "orange") + theme_bw() + 
  labs(x = "Date", y = "Exponential Growth Coefficient", title="On-Premises Counties\n Mobility 4 weeks forward",
       color = "Majority Teaching Method",
       caption = "Smoothing window set to every 4 weeks",
       subtitle = "Yellow area represents the fall semester (08/18 - 12/15)") + 
  theme(legend.position = "bottom")+
  geom_line(aes(x = DATE,y=cum_full_work/1000000),color="orange")+
  scale_y_continuous(,# Add a second axis and specify its features
  sec.axis = sec_axis(trans=~.,name="Cumulative Full-time Workers/Students",labels =c("-1"," ","0","25k","50k","75k")))+annotate("text",label="Exponential Growth Coefficient",x=as.Date("2020-09-21"),y=-0.010)+annotate("text",label="Cumulative Full-time Workers/Students",x=as.Date("2020-06-21"),y=0.040)+
  theme(legend.title = element_text(size=13),legend.text = element_text(size=12),axis.title = element_text(size=12),axis.text = element_text(size=12),title = element_text(size=12.9))+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) 
ggsave("time_series_on_premises.png",width = 7, height = 5)
```

```{r}
cases_slope_teach_agg%>%
  filter(major_teaching=="On Premises")%>%
  inner_join(cum_mobility%>%filter(major_teaching=="On Premises"),by=c("DATE"="shiftedDate"))%>%
  ggplot(aes(y = B,x=cum_full_work))+ 
  geom_point(na.rm = TRUE,colour = 'salmon') + 
  geom_smooth(method = "lm", formula = y~x, se = FALSE, na.rm = TRUE, show.legend = FALSE,colour = 'black') + 
  theme_minimal() + theme(legend.position = "bottom")+
  labs(x = "Cumulative Full-time Workers/Students", y = "Exponential Growth Coefficient", 
       title = "B ~ Cumulative Full-time Workers/Students 4 weeks forward\n On Premises")+theme(text = element_text(size=14),title = element_text(size=13))

#ggsave("lm_on_premises.jpg", width = 15, height = 6)
```


Plot2.2: Hybrid
```{r}
cases_slope_teach_agg%>%
  filter(major_teaching=="Hybrid")%>%
  inner_join(cum_mobility%>%filter(major_teaching=="Hybrid"),by=c("DATE"="shiftedDate"))%>%
  ggplot(aes(x = DATE))+
  geom_line(aes(y = B),size = 1,color="darkgreen") +
  geom_rect(data = cases_slope_teach_agg[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "orange") + theme_bw() + 
  labs(x = "Date", y = "Exponential Growth Coefficient", title="Hybrid Counties\n Mobility 4 weeks forward",
       color = "Majority Teaching Method",
       caption = "Smoothing window set to every 4 weeks",
       subtitle = "Yellow area represents the fall semester (08/18 - 12/15)") + 
  theme(legend.position = "bottom")+
  geom_line(aes(x = DATE,y=cum_full_work/1000000),color="green")+
  scale_y_continuous(# Add a second axis and specify its features
  sec.axis = sec_axis(trans=~.*10,name="Cumulative Full-time Workers/Students",labels = c(" "," ","0","200k","300k","400k","500k")))+
  annotate("text",label="Exponential Growth Coefficient",x=as.Date("2020-09-21"),y=0.06)+
  annotate("text",label="Cumulative Full-time Workers/Students",x=as.Date("2020-06-21"),y=0.40)+
  theme(legend.title = element_text(size=13),legend.text = element_text(size=12),axis.title = element_text(size=12),axis.text = element_text(size=12),title = element_text(size=12.9))+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) 

#ggsave("time_series_hybrid.png",width = 7, height = 5)
```

```{r}
cases_slope_teach_agg%>%
  filter(major_teaching=="Hybrid")%>%
  inner_join(cum_mobility%>%filter(major_teaching=="Hybrid"),by=c("DATE"="shiftedDate"))%>%
  ggplot(aes(y = B,x=cum_full_work))+ 
  geom_point(na.rm = TRUE,colour = 'darkgreen') + 
  geom_smooth(method = "lm", formula = y~x, se = FALSE, na.rm = TRUE, show.legend = FALSE,colour = 'black') + 
  theme_minimal() + theme(legend.position = "bottom")+
  labs(x = "Cumulative Full-time Workers/Students", y = "Exponential Growth Coefficient", 
       title = "B ~ Cumulative Full-time Workers/Students 4 weeks forward\n Hybrid")+theme(text = element_text(size=14),title = element_text(size=13))
  
```

Plot2.3: Online Only
```{r}
cases_slope_teach_agg%>%
  filter(major_teaching=="Online Only")%>%
  inner_join(cum_mobility%>%
               filter(major_teaching=="Online Only"),by=c("DATE"="shiftedDate"))%>%
  ggplot(aes(x = DATE))+
  geom_line(aes(y = B),size = 1,color="darkblue") +
  geom_rect(data = cases_slope_teach_agg[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "orange") + theme_bw() +
  labs(x = "Date", y = "Exponential Growth Coefficient", title="Online-only Counties\n Mobility 4 weeks forward",
       color = "Majority Teaching Method",
       caption = "Smoothing window set to every 4 weeks",
       subtitle = "Yellow area represents the fall semester (08/18 - 12/15)") + 
  theme(legend.position = "bottom")+
  geom_line(aes(x = DATE,y=cum_full_work/1000000),color="blue")+
  scale_y_continuous(# Add a second axis and specify its features
  sec.axis = sec_axis(trans=~.*10,name="Cumulative Full-time Workers/Students",labels = c(" ","0","100k","200k","300k","400k")))+
  annotate("text",label="Exponential Growth Coefficient",x=as.Date("2020-09-21"),y=0.06)+
  annotate("text",label="Cumulative Full-time Workers/Students",x=as.Date("2020-06-21"),y=0.40)+
  theme(legend.title = element_text(size=13),legend.text = element_text(size=12),axis.title = element_text(size=12),axis.text = element_text(size=12),title = element_text(size=12.9))+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) 

#ggsave("time_series_online_only.png",width = 7, height = 5)
```


```{r}
cases_slope_teach_agg%>%
  filter(major_teaching=="Online Only")%>%
  inner_join(cum_mobility%>%filter(major_teaching=="Online Only"),by=c("DATE"="shiftedDate"))%>%
  ggplot(aes(y = B,x=cum_full_work))+ 
  geom_point(na.rm = TRUE,colour = 'darkblue') + 
  geom_smooth(method = "lm", formula = y~x, se = FALSE, na.rm = TRUE, show.legend = FALSE,colour = 'black') + 
  theme_minimal() + theme(legend.position = "bottom")+
  labs(x = "Cumulative Full-time Workers/Students", y = "Exponential Growth Coefficient", 
       title = "B ~ Cumulative Full-time Workers/Students 4 weeks forward\n Online Only")+theme(text = element_text(size=14),title = element_text(size=13))
  
```
