---
title: "Interpretation"
author: "PHIGHT COVID"
date: "3/23/2021"
output: pdf_document
---

#Get Wranggling Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rstatix)
library(ggpubr)

bracket_county_policy_aggregate_wide <- read.csv("~/Desktop/PHIGHT/03:24/bracket_county_policy_aggregate_wide.csv")
bracket_county_policy_aggregate_wide$major_teaching <- as.factor(bracket_county_policy_aggregate_wide $major_teaching)

#we might want per 100,000 people? county pop 10k ~ 1000k
bracket_county_policy_aggregate_wide <- bracket_county_policy_aggregate_wide %>%
  mutate(deaths_increase_per_100k = 100*deaths_increase_per_1000)
```


# 1-Way ANOVA on Teaching Method (deaths)
```{r}
tm_bracket_deathprop.aov <- aov(deaths_increase_per_100k ~major_teaching,data = bracket_county_policy_aggregate_wide)

summary(tm_bracket_deathprop.aov)
tm_bracket_deathprop.aov$coefficients
```

We have an ANOVA that has detected a significant effect of the factor, which in this case is teaching method. We know this because the p-value < 0.05.

# 1-Way ANOVA on Teaching Method (cases)
```{r}
tm_bracket_caseprop.aov <-  aov(cases_increase_per_1000 ~major_teaching,data = bracket_county_policy_aggregate_wide)

summary(tm_bracket_caseprop.aov)
tm_bracket_caseprop.aov$coefficients
```

We have an ANOVA that has detected a significant effect of the factor, which in this case is teaching method. We know this because the p-value < 0.1. However, the ANOVA fails to detect the significance on the 0.05 level. 


We therefore conclude that teaching method is an important factor that has a real impact on the transmission of COVID. The impact, is more obvious reflected by deaths increase than cases increase. 


# Multiple Comparison (deaths)

We conduct statistical tests to help us understand the group difference using multiple comparisons. 


# Pairwise T-tests for multiple groups
```{r}
res.aov <- bracket_county_policy_aggregate_wide %>% anova_test(deaths_increase_per_100k ~major_teaching)
# Pairwise comparisons
pwc <- bracket_county_policy_aggregate_wide %>%
  pairwise_t_test(deaths_increase_per_100k ~major_teaching, p.adjust.method = "bonferroni")
pwc
```

```{r}
# Show adjusted p-values
pwc <- pwc %>% add_xy_position(x = "major_teaching")
ggboxplot(bracket_county_policy_aggregate_wide, x = "major_teaching", y = "deaths_increase_per_100k") +
  stat_pvalue_manual(pwc, label = "p.adj", tip.length = 0, step.increase = 0.1) +
  labs(
    subtitle = get_test_label(res.aov, detailed = TRUE),
    caption = get_pwc_label(pwc)
    )
```

```{r}
# Show significance levels
# Hide non-significant tests
ggboxplot(bracket_county_policy_aggregate_wide, x = "major_teaching", y = "deaths_increase_per_100k") +
  stat_pvalue_manual(pwc, hide.ns = TRUE, label = "p.adj.signif") +
  labs(
    subtitle = get_test_label(res.aov, detailed = TRUE),
    caption = get_pwc_label(pwc)
    )
```


* Tukey

The Tukey HSD procedure will run a pairwise comparison of all possible combinations of groups and test these pairs for significant differences between their means

```{r}
(tuk <- TukeyHSD(tm_bracket_deathprop.aov))
plot(tuk)
```

Significant: 
On Premises - Hybrid
On Premises - Online Only


Explanation:
There is no significant difference in deaths increase between Hybrid and Online Only

THERE IS A SIGNIFICANT DIFFERENCE in deaths increase between On Premises and Hybrid, On Premises and Online Only


Interpretation:
diff: mean (?should we increase to 100k)

If we switch from in person teaching to hybrid, we potentially save 32 lives per 100k people.
If we switch from in person teaching to online, we potentially save 47 lives per 100k people.

The CI looks big, should we say we can save up to 60 / 85 lives??

Another way to do this on a state level(%):

```{r}
tm_bracket_deathprop.aov1 <- aov(deaths_increase_prop ~major_teaching,data = bracket_county_policy_aggregate_wide)

summary(tm_bracket_deathprop.aov1)

(tuk1 <- TukeyHSD(tm_bracket_deathprop.aov1))
plot(tuk1)
```

```{r}
#state population
(p <- sum(bracket_county_policy_aggregate_wide$POPULATION))

#calculate how many lives saved
0.0003152*p
0.0004732*p

#current total deaths increase
(d <- sum(bracket_county_policy_aggregate_wide$deaths_increase, na.rm = TRUE))
```

```{r}
3696/d
5549/d
```






# Other Comparison Tests (tried)

```{r}
library(DescTools)
 
PostHocTest((tm_bracket_deathprop.aov), method = "duncan")

PostHocTest((tm_bracket_deathprop.aov), method = "scheffe")

DunnettTest(bracket_county_policy_aggregate_wide$deaths_increase_per_100k, bracket_county_policy_aggregate_wide $major_teaching, control = "On Premises")
```

Results pretty consistent in terms of estimates, p-values differ.


# Kruskal-Wallis (no assumptions)

```{r}
kruskal.test(bracket_county_policy_aggregate_wide$deaths_increase_per_100k, bracket_county_policy_aggregate_wide $major_teaching)
```


