---
title: "Step 1 Data Cleaning"
author: "Ziyan Zhu"
date: "3/15/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(mice)
library(visdat)
library(readxl)
```

# Ohio K12 Data
## Raw data

```{r}
oh_k12 <- read_excel("OH_K12.xlsx")
head(oh_k12)
```
```{r}
glimpse(oh_k12)
```



## Objectives in data cleaning


- Drop redundant columns

- Spot ill-posed data and potential wrong data entries

- Clean the structure for easier manipulation

- Remove observations with too much missing values

### Features we can drop:

- srcname: most school names is missing
- ncessch: only one single value
- city & physicalcity: we concern about county level data but we will keep physicalcity for impute missing values in county
- level_ is seemed to be correlated to level (1: elementary, 2: middle, 3:high)
- stusab/physicalstate: single value of OH
- schoolyear: single value of the same academic year 2020-2021
- district_nces & leaid are the same identifiers of school district, keep only leaid


```{r}
# Rename the columns with lower cases for easier reference
names(oh_k12) <- tolower(names(oh_k12))

# deselect redundant columns
oh_k12 <- oh_k12 %>% 
  select(-stusab, -srcname, -ncessch, -city, -level_,-schoolyear, -district_nces)
```

### transform data type

```{r}
oh_k12$enrollment <- as.numeric(oh_k12$enrollment)
oh_k12$enrollment_1 <- as.numeric(oh_k12$enrollment_1)
oh_k12$enrollment[oh_k12$enrollment==-999] <- NA
oh_k12[1:10,c('enrollment','enrollment_1')]
```

__Notes:__

Enrollment marks the number of students in a school
Enrollment_1 is the total number of students in its school district


```{r}
# columns with dates
any_date <- function(col){
  any(str_detect(col, "20..-"),na.rm = T)
}

# which are date columns?
oh_k12[ ,apply(oh_k12,2,any_date)]

# transform into Date obj 
oh_k12$date <- as.Date(oh_k12$date)
oh_k12$lastverifieddate <- as.Date(oh_k12$lastverifieddate)
oh_k12$opendategrouped <- as.Date(oh_k12$opendategrouped)
```


```{r}
# check the entries of columns storing strings
char_cols <- oh_k12[ ,apply(oh_k12,2,is.character)]
colnames(char_cols)

apply(char_cols[,-c(1:2,9,10,14,15,27,28,8,12)], 2, table)
```



### missing values

```{r}
# convert N to NA
oh_k12$gslo[oh_k12$gslo=='N'] <- NA
oh_k12$gshi[oh_k12$gshi=='N'] <- NA
oh_k12$openenroll[oh_k12$openenroll==2] <-NA
oh_k12$level[oh_k12$level=='N'] <- NA
```


```{r}
ini <- mice(oh_k12,maxit = 0)
# how many missing values for each column
ini$nmis
# check if there is patterns in missing values
vis_miss(oh_k12)
```


1. Impute missing values in county

```{r}
# create pairs of physical city and county names for reference
county_city <- oh_k12%>%
  group_by(county)%>%
  select(county,physicalcity)%>%
  unique()%>%
  na.omit(county)

county_city

# create id for rows with missing values in county
miss_county_id <- oh_k12%>%
  mutate(id = row_number())%>%
  filter(is.na(county))%>%
  select(id, schnam,county,physicalcity,multibdy)
miss_county_id 
# 58 rows with missing in county
# dim(impute_county_id)

# merge missing value df with  reference
impute_county <- miss_county_id%>%
  left_join(county_city,by = 'physicalcity')%>%
  rename(county = county.y)%>%
  select(id,schnam,county,physicalcity,multibdy)
impute_county

dim(impute_county)

# row number does not match -> one city matches with two county
impute_county%>%
  filter(duplicated(id))
```
__NOTES__:

Research shows that Dublin is a city in Franklin, Delaware, and Union counties in the U.S. state of Ohio, we will note impute the county for this city.

Clermont and Hamilton makes up Cincinnati city, will impute that part manually.

```{r}
# not impute dublin and cincinnati
impute_county_nam <- miss_county_id%>%
  filter(!physicalcity %in% c('Dublin','Cincinnati'))%>%
  left_join(county_city,by = 'physicalcity')%>%
  rename(county = county.y)%>%
  select(id,schnam,county,physicalcity,multibdy)

oh_k12[impute_county_nam$id,'county'] <- impute_county_nam$county
oh_k12[is.na(oh_k12$county),'county'] <- c(rep('CLERMONT',2),rep('FRANKLIN',3))
```

2. handle missing values

```{r}
oh_k12%>%
  filter(is.na(oh_k12$teachingmethod))

oh_k12 <- oh_k12[!is.na(oh_k12$teachingmethod),]

oh_k12[is.na(oh_k12$enrollment),]%>%
  filter(studentmaskpolicy %in% c('Required for high school students only',"Required for middle/high school students only"))
```


3. aggregate enrollment

```{r}
oh_k12%>%
  group_by(leaid)%>%
  mutate(total_enrollment = sum(enrollment,na.rm = T))%>%
  select(county,leaid,enrollment,total_enrollment,enrollment_1)%>%
  filter(total_enrollment>enrollment_1)


check_enroll <- oh_k12%>%
  select(county,leaid,enrollment,enrollment_1)%>%
  group_by(county,leaid,enrollment_1)%>%
  summarise(leaid_enrollment = sum(enrollment,na.rm = T))

check_enroll <- check_enroll%>%
  mutate(district_enroll = ifelse(leaid_enrollment>enrollment_1,leaid_enrollment, enrollment_1))%>%
  select(-enrollment_1)


oh_k12 <- check_enroll%>%
  group_by(county)%>%
  mutate(county_enroll = sum(district_enroll))%>%
  full_join(oh_k12,by = c('county','leaid'))%>%
  select(-leaid_enrollment,enrollment_1)


```




```{r}
#export the data
oh_k12$lastverifieddate <- as.character(oh_k12$lastverifieddate)
oh_k12$opendategrouped <- as.character(oh_k12$opendategrouped)
write_excel_csv(oh_k12,"OH_K12_clean.csv")
```

