---
title: "Pairwise analysis on bracket"
author: "PHIGHT COVID"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Get Wrangled data

```{r, message = FALSE, warning=FALSE}
source('step3_bracket_data.R')
bracket_county_policy_wide$major_teaching <- as.factor(bracket_county_policy_wide$major_teaching)
```


# One Way ANOVA on Teaching Method
```{r}
tm_deathprop.aov <- aov(death_prop ~ major_teaching,data = county_policy_wide)
summary(tm_deathprop.aov)

tm_bracket_deathprop.aov <- aov(deaths_increase_prop ~major_teaching,data = bracket_county_policy_wide)

summary(tm_bracket_deathprop.aov)


tm_bracket_caseprop.aov <-  aov(cases_increase_prop ~major_teaching,data = bracket_county_policy_wide)
summary(tm_bracket_caseprop.aov )
```

# Pairwise mean comparison on bracket data

Because the ANOVA test shows insignificance between teaching methods in ALL DATA. We only perform pairwise tests on bracket data.

## Deaths

The differences between "Online Only - On Premises" and "Other - On Premises" are slightly significant in the duncan test. In the Dunnett test, the "Online Only - On Premises" difference is also slightly significant. 

```{r}
#library(PMCMRplus)
#library(DescTools)
 
#PostHocTest((tm_deathprop_s.aov), method = "duncan")

#PostHocTest((tm_deathprop_s.aov), method = "scheffe")

#DunnettTest(county_data$death_prop_s, county_data$teachingmethod, control = "Online Only")
```


# One Way Anova Against Online Only

```{r}
long_isonline$is_online <- as.factor(long_isonline$is_online)

tm_deathprop.aov <- aov(death_prop ~ is_online,data = long_isonline)
summary(tm_deathprop.aov)

isonline_fit <- lm(death_prop~is_online,long_isonline)
summary(isonline_fit)
```


```{r}
long_notonline_mask <- long_isonline_mask%>%
  filter(is_online=='Not Online Only')

summary(aov(death_prop ~ (prop_student_mask>0.8),data = long_notonline_mask))

summary( aov(death_prop ~ (prop_student_mask>0.98),data = long_notonline_mask))


```



```{r}
long_bracket_notonline_mask <- long_bracket_isonline_mask%>%
  filter(is_online=='Not Online Only')

summary(aov(deaths_increase_prop ~ (prop_student_mask>0.8),data = long_bracket_notonline_mask))

summary( aov(deaths_increase_prop ~ (prop_student_mask>0.94),data = long_bracket_notonline_mask))

```


```{r}

long_teaching_mask%>%
  group_by(teachingmethod)%>%
  summarise(sample_size = n())


hybrid_cutoff <- function(cutoff){
  aovs <- summary(aov(death_prop ~ (prop_student_mask>cutoff),data = long_teaching_mask%>%filter(teachingmethod=='Hybrid')))
 pval <- aovs[[1]]$`Pr(>F)`[1]
 return(pval)
}

pvals_hybrid <- unlist(lapply(seq(0.4,0.99,by=0.01), hybrid_cutoff))

data.frame(cutoffs=seq(0.4,0.99,by=0.01),pvals_hybrid)%>%
ggplot(aes(x = cutoffs,y = pvals_hybrid))+geom_point(size = 3,color='salmon3')+theme_minimal()+geom_abline(intercept = 0.05,slope = 0)+labs(x = 'Cutoff for Proportion of student wearing mask',y = 'P-value for On Hybrid')+theme(title=element_text(size=15))



premise_cutoff <- function(cutoff){
  aovs <- summary(aov(death_prop ~ (prop_student_mask>cutoff),data = long_teaching_mask%>%filter(teachingmethod=='On Premises')))
 pval <- aovs[[1]]$`Pr(>F)`[1]
 return(pval)
}

pvals_premise <- unlist(lapply(seq(0.4,0.99,by=0.01), premise_cutoff))

data.frame(cutoffs=seq(0.4,0.99,by=0.01),pvals_premise)%>%
ggplot(aes(x = cutoffs,y = pvals_premise))+geom_point(size = 3,color='salmon3')+geom_abline(intercept = 0.05,slope = 0)+labs(x = 'Cutoff for Proportion of student wearing mask',y = 'P-value for On Premises')+theme_minimal()+theme(title=element_text(size=15))


#summary( aov(death_prop ~ (prop_student_mask>0.5),data = long_teaching_mask%>%filter(teachingmethod=='Online Only')))
```
```{r}

hybrid_cutoff <- function(cutoff){
aovs <- summary( aov(deaths_increase_prop ~ (prop_student_mask>cutoff),data = long_bracket_teaching_mask%>%filter(teachingmethod=='Hybrid')))
 pval <- aovs[[1]]$`Pr(>F)`[1]
 return(pval)
}

pvals_hybrid <- unlist(lapply(seq(0.4,0.99,by=0.01), hybrid_cutoff))

data.frame(cutoffs=seq(0.4,0.99,by=0.01),pvals_hybrid)%>%
ggplot(aes(x = cutoffs,y = pvals_hybrid))+geom_point(size = 3,color='salmon3')+theme_minimal()+geom_abline(intercept = 0.05,slope = 0)+labs(x = 'Cutoff for Proportion of student wearing mask',y = 'P-value for On Hybrid')+theme(title=element_text(size=15))


premise_cutoff <- function(cutoff){
  aovs <- summary(aov(deaths_increase_prop ~ (prop_student_mask>cutoff),data = long_bracket_teaching_mask%>%filter(teachingmethod=='On Premises')))
 pval <- aovs[[1]]$`Pr(>F)`[1]
 return(pval)
}

pvals_premise <- unlist(lapply(seq(0.4,0.99,by=0.01), premise_cutoff))

data.frame(cutoffs=seq(0.4,0.99,by=0.01),pvals_premise)%>%
ggplot(aes(x = cutoffs,y = pvals_premise))+geom_point(size = 3,color='salmon3')+geom_abline(intercept = 0.05,slope = 0)+labs(x = 'Cutoff for Proportion of student wearing mask',y = 'P-value for On Premises')+theme_minimal()+theme(title=element_text(size=15))

seq(0.4,0.99,by=0.01)[which(pvals_premise<0.05)]
```

# Linear Regression Attempts


```{r}

library(corrplot)

county_policy_wide%>%
  select(On_Premises,Online_Only)%>%
  drop_na()%>%
  cor()





plot(x= county_policy_wide$Online_Only,y=county_policy_wide$death_prop,col=as.factor(county_policy_wide$major_teaching))

cor(county_policy_wide$On_Premises,county_policy_wide$Online_Only,na.rm=TRUE)

summary(lm(log(death_prop)~On_Premises,data = county_policy_wide))

summary(lm(log(death_prop)~Online_Only,data = county_policy_wide))

teaching_fit <- lm(log(death_prop)~deathBefore+Online_Only+On_Premises,data = county_policy_wide)
summary(teaching_fit)



```

```{r}
teaching_fit <- lm(log(death_prop)~Hybrid+Online_Only+On_Premises,data = county_policy_wide)
summary(teaching_fit)
```



```{r}
mask_fit <- lm(log(death_prop)~Hybrid*Required_for_all_students+Online_Only*Required_for_all_students+On_Premises*Required_for_all_students,data = county_policy_wide)
summary(mask_fit)
```

```{r}
wide_teaching_enroll%>%
  filter(On_Premises>0.6)%>%
  select(county,On_Premises,county_enroll)%>%
  right_join(cases,by=c('county'='COUNTY'))%>%
  right_join(OH_K12%>%select(county,opendategrouped),by='county')%>%
  group_by(DATE)%>%
  summarise(AVG_CUMDEATHS = mean(CUMDEATHS,na.rm = T))%>%
  ggplot(aes(x=DATE,y=AVG_CUMDEATHS))+geom_point()
  
```


