---
title: "EDA on County Policy Data"
author: "CMU MSP Team"
date: "3/20/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


# Documentation

- For the first part of the EDA, we focus on the latest cumulative deaths on __2021/02/22__

- To measure to what extent each policy was carried out in a county:
  - since we find that all schools in the same district have same public health strategies
  - for each county, we weight the count of policies by #students enrolled in a district/#students enrolled in a county

- For the second part of the EDA, we analyze the covid data on the schools’ open days only       (From county-level open_date_grouped to 2020/12/15)

### Naming Conventions

- OH_K12: cleaned ohio k12 data
  - #enrollment_1: number of students enrolled in each district

- cases: ohio county level cases data

- county_w_dates

- county_policy: policy counts proportion to #enrollment at county-level

- county_data: policy data on schools' open days

# Check Data 

### Read data

```{r, message = FALSE}
# libraries
library(tidyverse)
library(readxl)
library(ggplot2)
library(reshape2)
library(wesanderson)
pal <- wes_palette(name = "GrandBudapest1", type = "discrete")
# read in data
#source("step2_data_wrangle.R")
source("step3_bracket_data.R")
```


### Check data quality

```{r}
# county-wise total deaths = sum of new deaths until Feb.2021
total_deaths_by_county <- cases %>%
  group_by(COUNTY) %>%
  summarise(total_deaths = sum(NEWDEATHS, na.rm = TRUE), .groups = "drop")

# latest collection date
tail(sort(unique(cases$DATE)))
# latest cumulative death
latest_cumdeaths <- cases%>%
  filter(DATE == '2021-02-22')%>%
  select(COUNTY, CUMDEATHS)%>%
  unique()

# check if the latest death matches with sum of new deaths
sum(total_deaths_by_county$total_deaths != latest_cumdeaths$CUMDEATHS)
cbind(total_deaths_by_county,latest_cumdeaths)%>%head(3)
```

### generate is online method

```{r}
isonline_enroll <- teachingmethod_enroll%>%
  filter(teachingmethod != 'Other'&teachingmethod != 'Pending'&teachingmethod != 'Unknown')%>%
  mutate(is_online = ifelse(teachingmethod == "Online Only","Online Only","Not Online Only"))%>%
  group_by(county,is_online) %>%
  mutate(prop_online_only = sum(prop_teachingmethod))%>%
  group_by(county)%>%
  slice(which.max(prop_online_only))%>%
  select(county,is_online,prop_online_only)
```


# Marginal Distribution of Number of students enrollents by Each Methods

```{r}
long_teaching%>%
  filter(teachingmethod %in% c('Online Only','On Premises','Hybrid'))%>%
  group_by(county)%>%
  slice(which.max(prop_teachingmethod))%>%
  ggplot(aes(x=prop_teachingmethod,group=teachingmethod,fill=teachingmethod))+geom_bar(stat = 'count',width=0.03,position = "stack")+labs(x = "Percentage of Students",y='Count',fill='Majority Teaching Method',title='Conditional Distribution of Percentages of Students',subtitle = 'by Most Common Teaching Method')+theme_minimal()+theme(text = element_text(size = 15),title = element_text(size=14),legend.position = 'bottom')+facet_wrap(~teachingmethod)+scale_fill_manual(values=pal)
  

studentmask_enroll%>%
  filter(studentmaskpolicy=='Required for all students')%>%
  ggplot(aes(y= prop_student_mask))+geom_bar(stat = 'count',width=0.03,fill='salmon3')+labs(x = "Count",y="Percentage of students wear mask")+coord_flip()+theme_minimal()+theme(axis.text = element_text(size = 16),axis.title = element_text(size=18))

studentmask_enroll%>%
  filter(studentmaskpolicy=='Not required')%>%
  ggplot(aes(y= prop_student_mask))+geom_bar(stat = 'count',width=0.03)+labs(x = "Count",y="Percentage of students not wear mask")+coord_flip()+theme_minimal()+theme(title = element_text(size=16),text=element_text(size = 15))


staffmask_enroll%>%
   filter(staffmaskpolicy=='Required for all staff')%>%
  ggplot(aes(y= prop_staff_mask))+geom_bar(stat = 'count',width=0.03,fill='salmon3')+labs(x = "Count",y="Percentage of staffs wear mask")+coord_flip()+theme_minimal()+theme(axis.text = element_text(size = 16),axis.title = element_text(size=18))
```

# Conditional Distribution of Number of students enrollents by Each Methods

```{r}
options(scipen=10000)
teachingmethod_enroll%>%
  group_by(teachingmethod)%>%
  summarise(total_teaching= sum(total_teachingmethod))%>%
  ggplot(aes(x= reorder(teachingmethod,total_teaching),y = total_teaching))+geom_bar(stat = 'identity',fill='salmon3',width = 0.8)+labs(y = "Number of Enrollments",x='Teaching Method',title='Distribution of Student Enrollments in Ohio by Teaching Method')+theme_minimal()+theme(axis.text = element_text(size = 11),title = element_text(size=12))+coord_flip()



OH_K12%>%
  distinct(county,schnam,leaid,district_enroll,studentmaskpolicy)%>%
  group_by(studentmaskpolicy)%>%
  count()
  
  summarise(n_total = sum(district_enroll))%>%
  group_by(county)%>%
  slice(which.max(n_total))%>%
  group_by(studentmaskpolicy)%>%
  count()


studentmask_enroll%>%
   group_by(studentmaskpolicy)%>%
  summarise(total_studentmask= sum(n_studentmask))
    ggplot(aes(x= reorder(studentmaskpolicy,total_studentmask),y=total_studentmask))+geom_bar(stat = 'identity',fill='salmon3')+labs(y = "Total Enrollments",x="Student Wearing Mask Proportions")+coord_flip()+theme_minimal()+theme(axis.text = element_text(size = 13),axis.title = element_text(size=16))

304798+134644	
    
staffmask_enroll%>%
  group_by(staffmaskpolicy)%>%
  summarise(total_staffmask= sum(n_staffmask))%>%
   ggplot(aes(x= reorder(staffmaskpolicy,total_staffmask),y = total_staffmask))+geom_bar(stat = 'identity')+labs(x='Staff Wearing Mask',y = "Staff Wearing Mask Proportions",subtitle = "We assume the number of staff proportionate to student enrollments")+coord_flip()+theme_minimal()+theme(axis.text = element_text(size = 15),title = element_text(size=16))

```

```{r}
ohioage <- read_excel("OhiobyAge.xlsx")
ohioage
ohioage[,1:11]%>%
  melt(id.var = c("Date","Area","Metric"),value.name = "Cases",variable.name='Age_Group')%>%
  filter(Age_Group %in% c('00_19','70_79','80+'))%>%
  ggplot(aes(x= Date,y = Cases,group=Age_Group,color = Age_Group))+geom_line()+theme_minimal()+theme(axis.title = element_text(size=14),axis.text = element_text(size=13))
```





# Conditional Distribution of Death Proportions

### Marginal Distribution of Death v.s. Population

```{r}
# look at influential points, size = total_county_enroll
county_policy_wide%>%
  ggplot(aes(x = log(POPULATION), y = log(CUMDEATHS), size=county_enroll)) + 
  geom_point(na.rm = TRUE,colour = 'salmon3') + 
  geom_smooth(method = "lm", formula = y~x, se = FALSE, na.rm = TRUE, show.legend = FALSE,colour = 'black') + 
  theme_minimal() + theme(legend.position = "bottom")+
  labs(x = "Log Population", y = "Log Total Deaths", 
       size = "Number of Enrollments", title = "Death against Population")+theme(text = element_text(size=14),title = element_text(size=16))

ggsave("DeathVSPopulation.jpg", height = 5, width = 7)
```

```{r}
county_policy_wide %>%
  ggplot(aes(x = death_prop)) + 
  geom_histogram(bins = 25, na.rm = TRUE,fill = 'salmon3') + 
  theme_minimal() + 
  labs(x = "Population Death Proportion", y = "Frequency", 
       title = "Marginal Distribution of Death Proportion")+theme(text = element_text(size=14),title = element_text(size=16))

ggsave("MargDeathProp.jpg", height = 5, width = 7)
```


### Conditional distribution of death proportions by Teaching Method

```{r}
county_policy_wide %>%
  drop_na()%>%
  ggplot(aes(y = death_prop, x = major_teaching, fill = major_teaching == 'Online Only')) + 
  geom_boxplot() + scale_fill_manual(values = pal[c(1,4)])+
  theme_minimal() + 
  theme(axis.text.x = element_text(size = 14),title = element_text(size = 14),legend.position = "bottom")+
  labs(y = "Population Death Proportion", 
       title = "Distribution of Death Proportion by Teaching Method",
       x = "Majority Teaching Method", fill = 'Online Teaching')

ggsave("DeathProp_teachingmethod.jpg", height = 5, width = 7)

bracket_county_policy_wide %>%
  drop_na()%>%
  filter()%>%
  ggplot(aes(y = deaths_increase_prop, x = major_teaching, fill = major_teaching == 'Online Only')) + 
  geom_boxplot() + scale_fill_manual(values = pal[c(1,4)])+
  theme_minimal() + 
  theme(axis.text.x = element_text(size = 14),title = element_text(size = 14),legend.position = "bottom")+
  labs(y = "Population Death Increase Proportion", 
       title = "Distribution of Death Increase Proportion by Teaching Method", subtitle = '*On bracket data',
       x = "Majority Teaching Method", fill = 'Online Teaching')
ggsave("DeathProp_bracket_teachingmethod.jpg", height = 5, width = 7)
```

```{r}
county_policy_wide%>%
  drop_na()%>%
  ggplot(aes(x = log(POPULATION),y = death_prop, color = major_teaching)) + geom_point(alpha = 0.5,size=2.5)+ 
  geom_smooth(method = "lm", formula = y~x,se = FALSE, show.legend = FALSE) + 
  theme_minimal() + 
  theme(axis.text = element_text(size = 11),title = element_text(size = 14),legend.text = element_text(size = 13),legend.position = "bottom")+
  labs(x = "Log Population", y = "Population Death Proportion", 
       color = "Teaching Method")

ggsave("DeathProp_onlineonly.jpg", height = 5, width = 7)

bracket_county_policy_wide%>%
  drop_na()%>%
  ggplot(aes(x = log(POPULATION),y = deaths_increase_prop, color = major_teaching)) + geom_point(alpha = 0.5,size=2.5)+ 
  geom_smooth(method = "lm", formula = y~x,se = FALSE, show.legend = FALSE) + 
  theme_minimal() + 
  theme(axis.text = element_text(size = 11),title = element_text(size = 14),legend.text = element_text(size = 13),legend.position = "bottom")+
  labs(x = "Log Population", y = "Population Death Increase Proportion", 
       color = "Teaching Method")
```


### Conditional distribution of death proportions by Online Teaching

```{r}

county_policy_wide %>%
  right_join(isonline_enroll,by='county')%>%
  ggplot(aes(y = death_prop, x = prop_online_only, fill= is_online)) + 
  geom_boxplot() + 
  theme_minimal() + 
   theme(axis.text.x = element_text(size = 11),title = element_text(size = 14),legend.position = "bottom")+
  labs(y = "Population Death Proportion", 
       x = "Proportion of students complied",
       title = "Distribution of Death Proportion by Online Method",
       fill = "Majority Teaching Method")

ggsave("DeathProp_online.jpg", height = 5, width = 7)
```

### Conditional distribution of death proportions by Mask Wearing

```{r}

long_studentmask %>%
  filter(!studentmaskpolicy%in%c('Pending','Unknown'))%>%
  group_by(county)%>%
  slice(which.max(prop_student_mask))%>%
  ggplot(aes(y = death_prop,x = (prop_student_mask>0.8),fill = (prop_student_mask>0.8))) +
  geom_boxplot() + scale_fill_manual(values = pal[c(1,4)])+
  theme_minimal() + 
  labs(y = "Population Death Proportion", x='Prop of Student Wearing Mask > 80%', title = "Distribution of Death Proportion by Student Wearing Mask",subtitle = "* Majority of students are required to wear mask for all counties",fill = 'Over 80%')+theme(legend.position = "bottom")

ggsave("DeathProp_student_mask.jpg", height = 5, width = 7)


long_bracket_isonline_mask%>%
  drop_na()%>%
  mutate(over_mask = prop_student_mask>0.8)%>%
  ggplot(aes(y = deaths_increase_prop,x=prop_student_mask,group = over_mask,fill = over_mask)) +
  geom_point(alpha=0.5,size=2,color='grey')+
  geom_boxplot() + scale_fill_manual(values = pal[c(1,4)])+
  theme_minimal() + facet_wrap(~is_online)+
  labs(y = "Population Death Increase Proportion", x='Prop of Student Wearing Mask > 80%', title = "Distribution of Death Increase Proportion \nby Mask Wearing for In-person Counties ",subtitle = "* Majority of students are required to wear mask for all counties",fill = 'Over 80%')+theme(legend.position = "bottom",text = element_text(size=14))

long_bracket_isonline_mask%>%
  drop_na()%>%
  mutate(over_mask = prop_student_mask>0.5)%>%
  ggplot(aes(y = deaths_increase_prop,x=prop_student_mask,group = over_mask,fill = over_mask)) +
  geom_point(alpha=0.5,size=2,color='grey')+
  geom_boxplot() + scale_fill_manual(values = pal[c(1,4)])+
  theme_minimal() + facet_wrap(~is_online)+
  labs(y = "Population Death Increase Proportion", x='Prop of Student Wearing Mask > 50%', title = "Distribution of Death Increase Proportion \nby Mask Wearing for In-person Counties ",subtitle = "* Majority of students are required to wear mask for all counties",fill = 'Over 80%')+theme(legend.position = "bottom",text = element_text(size=14))

```

```{r}
long_teaching_mask%>%
  drop_na()%>%
  filter(teachingmethod!='Online Only')%>%
  mutate(over_mask = prop_student_mask>0.8)%>%
  ggplot(aes(y = death_prop,x=prop_student_mask,group = over_mask,fill = over_mask)) +
  geom_point(alpha=0.5,size=2,color='grey')+
  geom_boxplot() + scale_fill_manual(values = pal[c(1,4)])+
  theme_minimal() + facet_wrap(~teachingmethod)+
  labs(y = "Population Death Proportion", x='Prop of Student Wearing Mask > 80%', title = "Distribution of Death Proportion \nby Mask Wearing for In-person Counties ",subtitle = "* Majority of students are required to wear mask for all counties",fill = 'Over 80%')+theme(legend.position = "bottom",text = element_text(size=14))


long_bracket_teaching_mask%>%
  drop_na()%>%
  filter(teachingmethod!='Online Only')%>%
  mutate(over_mask = prop_student_mask>0.8)%>%
  ggplot(aes(y = deaths_increase_prop,x=prop_student_mask,group = over_mask,fill = over_mask)) +
  geom_point(alpha=0.5,size=2,color='grey')+
  geom_boxplot() + scale_fill_manual(values = pal[c(1,4)])+
  theme_minimal() + facet_wrap(~teachingmethod)+
  labs(y = "Population Death Increase Proportion", x='Prop of Student Wearing Mask > 80%', title = "Distribution of Death Increase Proportion \nby Mask Wearing for In-person Counties ",subtitle = "* Majority of students are required to wear mask for all counties",fill = 'Over 80%')+theme(legend.position = "bottom",text = element_text(size=14))
```



```{r}
isonline_enroll%>%
  left_join(death_prop,by=c('county'='COUNTY'))%>%
  ggplot(aes(x = is_online, y = death_prop)) +
  geom_boxplot(na.rm = TRUE) +
  theme_minimal() + 
  labs(x = "Student Wear Mask", y = "Population Death Proportion",
       title = "Conditional Distribution of Death Proportion b Online Only")

ggsave("DeathProp_onlineonly.jpg", height = 5, width = 7)
```

```{r}

long_staff %>%
  filter(!staffmaskpolicy%in%c('Pending','Unknown'))%>%
  group_by(county)%>%
  slice(which.max(prop_staff_mask))%>%
  ggplot(aes(y = death_prop,x = (prop_staff_mask>0.8),fill = (prop_staff_mask>0.8))) +
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "Population Death Proportion", x='Prop of Staff Wearing Mask > 80%', title = "Distribution of Death Proportion by Staff Wearing Mask",subtitle = "* Majority of staffs are required to wear mask for all counties")+theme(legend.position = "")

ggsave("DeathProp_staff_mask.jpg", height = 5, width = 7)

```





