---
title: "Technical Appendix"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,warning=FALSE,error = FALSE)
```

# Notes for appendix

The Appendix still needs more comments and more details.

For the draft, we only contain the code for plots in our IDMRD paper.

# Appendix 1: Map

```{r,warning=FALSE,message=FALSE}
Sys.setlocale("LC_TIME", "English")
library(ggrepel)
library(cowplot)
library(sp)
source("step2_data_wrangle.R")
# color blind friendly Palette
library(ggthemes)
col_theme <- c("Hybrid"="#009E73","On Premises"="#D55E00","Online Only"="#0072B2")
## plot theme
grid_theme <- theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    legend.key = element_blank(),
    panel.background = element_blank(),
    legend.box="vertical", legend.margin=margin())
team_theme <- grid_theme+
  theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        axis.text = element_text(size=13),
        title=element_text(size=13),
        strip.text.x = element_text(size = 10, face = "bold.italic"))

```

## Teaching method, Population and Enrollment
```{r}
ohio_map <- map_data("county") %>%subset(region=="ohio")%>%
  mutate(county=toupper(subregion))%>%select(long,lat,county,group)
# create map plots
wide_teaching_enroll%>%
  left_join(ohio_map,by='county')%>%
  mutate(Online_Only= Online_Only*100)%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = Online_Only), color = "gray") + 
  coord_fixed(1.3) + theme_map() +
  scale_fill_distiller(palette = "OrRd",direction = 1)+
  labs(fill='% Online Only')+
  theme(legend.text = element_text(size=20),legend.title = element_text(size=20))
# create map plots
wide_teaching_enroll%>%
  left_join(ohio_map,by='county')%>%
  mutate(On_Premises= On_Premises*100)%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = On_Premises), color = "gray") + 
  coord_fixed(1.3) + theme_map() +
  scale_fill_distiller(palette = "OrRd",direction = 1)+
  labs(fill='% On Premises')+
  theme(legend.text = element_text(size=20),legend.title = element_text(size=20))
# create map plots for population
wide_teaching_enroll%>%
  left_join(ohio_map,by='county')%>%
  mutate(Hybrid= Hybrid*100)%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = Hybrid), color = "gray") + 
  coord_fixed(1.3) + 
  theme_map() +
  scale_fill_distiller(palette = "OrRd",direction = 1)+
  labs(fill='% Hybrid')+
  theme(legend.text = element_text(size=20),legend.title = element_text(size=20))
# create map plots
cases%>%
  distinct(COUNTY,POPULATION)%>%
  left_join(ohio_map,by=c('COUNTY'='county'))%>%
  mutate(population = POPULATION/1000)%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = population), color = "gray") + 
  coord_fixed(1.3) + theme_map() +
  scale_fill_distiller(palette = "OrRd",direction = 1)+
  labs(fill='Population/1000')+
  theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "bottom",
        legend.key.size = unit(2,"lines"))
# create map plots
teachingmethod_enroll%>%
  distinct(county,county_enroll)%>%
  left_join(ohio_map,by=c('county'))%>%
  mutate(county_enroll = county_enroll/1000)%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = county_enroll), color = "gray") + 
  coord_fixed(1.3) + theme_map() +
  scale_fill_distiller(palette = "OrRd",direction = 1)+
  labs(fill='Enrollment/1000')+
  theme(legend.text = element_text(size=12),legend.title = element_text(size=12),
        legend.position = "bottom",legend.key.size = unit(2,"lines"))
```

```{r}
wide_teaching_enroll%>%
  left_join(ohio_map,by='county')%>%
  mutate(On_Premises= On_Premises*100)%>%
  ggplot() + geom_polygon(aes(x = long, y = lat, group = group, fill = as.factor(major_teaching)), color = "white",alpha=0.9) + 
  coord_fixed(1.3) + theme_map() + 
  scale_fill_manual(values=col_theme)+ 
  labs(fill='Majority teaching posture')+
  theme(legend.position = "bottom",legend.text = element_text(size=14),legend.title = element_text(size=14))
```


## Covid deaths during fall semester and death proportion during fall semester

```{r}
getLabelPoint <- # Returns a county-named list of label points
function(county) {Polygon(county[c('long', 'lat')])@labpt}
centroids = by(ohio_map, ohio_map$county, getLabelPoint)# Returns list
centroids2 <- do.call("rbind.data.frame", centroids)# Convert to Data Frame
centroids2$county = str_to_title(rownames(centroids))
names(centroids2) <- c('clong', 'clat', "county") # Appropriate Header
```

```{r}
death_prop%>%
  left_join(ohio_map,by=c("COUNTY"='county'))%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group=group,fill = CUMDEATHS), color = "gray")+
  coord_fixed(1.3) + theme_map() +
  scale_fill_distiller(palette = "OrRd",direction = 1)+
  labs(fill='Cumulative Deaths \nuntil 2021-02-22')+
  theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),legend.position = "bottom",
        legend.key.size = unit(2,"lines"))
ggsave("cumdeath.png",width = 5, height = 5)
death_prop%>%
  left_join(ohio_map,by=c("COUNTY"='county'))%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group=group,fill = death_per_1000), 
               color = "gray") + 
  coord_fixed(1.3) + theme_map() +
  scale_fill_distiller(palette = "OrRd",direction = 1)+
  labs(fill='Deaths per 1000 people \nuntil 2021-02-22')+
  theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "bottom",
        legend.key.size = unit(2,"lines"))
ggsave("deathprop.png",width = 5, height = 5)
```


# Appendix 2: Death Incidence

## Data Process

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
require(scales)
library(readxl)
cases_by_age <- read_excel("OhiobyAge.xlsx")
rolling_age_cases <- cases_by_age %>%
  mutate(youth_prop_roll = zoo::rollmean(`00_19/total(%)`, k = 7, fill = NA),
         all_roll = zoo::rollmean(`00_80+`, k = 7, fill = NA))
colors <- c("Total Daily Cases" = "black",
            "0-19 Age / Total Cases (%)" = "gray")
coeff <- 200
cases_by_age_long <- cases_by_age %>%
  gather(age_group, percent_cases, 
         `00_19/total(%)`:`80+/total(%)`,
         factor_key=TRUE) %>%
  group_by(age_group) %>%
  mutate(roll_percent_cases= zoo::rollmean(percent_cases, k = 7, fill = NA))
county_policy_wide$major_teaching <- factor(county_policy_wide$major_teaching,
                                            levels = c("On Premises","Hybrid","Online Only"))
# see when the intesection happens
date.intercept <- as.Date("2020-11-24")
# add 95% confidence bans
confidence_level <- .95
z_cl <- qnorm(confidence_level)
# case_policy_wide
case_policy_wide <- cases %>%
  left_join(county_policy_wide[,c("county","major_teaching","Online_Only","Hybrid","On_Premises")],
            by = c("COUNTY" = "county")) %>%
  mutate(death_prop = CUMDEATHS/POPULATION)
opendate_cases <- case_policy_wide%>%
  inner_join(major_reopening%>%select(COUNTY,major_opendate),by=c('COUNTY'))
# Box Plots in Fall semester
library(PMCMRplus)
require(DescTools)
fall_cases <- opendate_cases %>%
  filter(DATE >= major_opendate & DATE <= as.Date("2020/12/15")) %>%
  group_by(COUNTY) %>%
  arrange(DATE) %>%
  filter(row_number()==1 | row_number()==n()) %>%
  mutate(death_incidence = diff(CUMDEATHS),
         death_incidence_per_1000 = death_incidence*1000/POPULATION) %>%
  distinct(COUNTY,POPULATION,major_teaching,
           death_incidence,death_incidence_per_1000) 
fall_major_teaching.aov <- aov(death_incidence_per_1000 ~ major_teaching,
                               data = fall_cases)
summary(fall_major_teaching.aov) # p-value of .012
stat.test <- PostHocTest(fall_major_teaching.aov, method = "duncan")$major_teaching%>%
  as.data.frame()%>%
  rownames_to_column("group") %>%
  separate(group,"-", into = c("group1","group2")) %>%
  mutate(pval = round(pval,3),
         p = case_when(pval <= .01~ "**",
                       pval <= .05 ~ "*",
                       TRUE ~ "NS"))%>%
  select(group1, group2, pval, p)
library(ggpubr)
```

## Death Prop Over Time by the Majority teaching posture
```{r}
case_policy_wide%>%
  group_by(DATE, major_teaching) %>%
  drop_na(major_teaching)%>%
  summarise(total_deaths = sum(CUMDEATHS),
            total_pop = sum(POPULATION),
            death_prop = total_deaths/total_pop,
            death_prop_upper = death_prop + z_cl*sqrt(death_prop*(1 - death_prop)/total_pop),
            death_prop_lower = death_prop - z_cl*sqrt(death_prop*(1 - death_prop)/total_pop),
            .groups = "drop") %>%
  ggplot(aes(x = DATE, y = death_prop*1000, group = major_teaching))+
  geom_rect(data=case_policy_wide[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "#E69F00") + 
  geom_line(aes(color = major_teaching),size = 1, alpha = .8) + 
  geom_ribbon(aes(ymin = 1000*death_prop_lower, ymax = 1000*death_prop_upper,
                  fill= major_teaching), 
              alpha = .3, show.legend = F)+ 
  geom_vline(xintercept = date.intercept, linetype = "dashed") + 
  annotate("text",x = date.intercept,y = 1.5,
           label = date.intercept,
           hjust = 1.1) + 
  theme_bw() + 
  ggtitle("Death Incidences Increase Faster for Red Counties ")+
  labs(x = "Date", y = "Cumulative Death Incidence / 1,000 people",
       subtitle = "Yellow area represents Fall Semester",
       color = "Majority teaching posture") +team_theme+ scale_colour_manual(values=col_theme)+scale_fill_manual(values=col_theme)
```


```{r}
ratio <- cases%>%
  left_join(wide_teaching_enroll,by=c("COUNTY"="county"))%>%
  drop_na(major_teaching)%>%
  group_by(major_teaching,DATE)%>%
  summarise(ratios = sum(CUMDEATHS)/sum(CUMCONFIRMED))

ratio%>%
  ggplot(aes(x=DATE,y=ratios,color=major_teaching))+geom_line(size=1)+
  geom_rect(data = ratio[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "orange") + 
  geom_vline(xintercept = as.Date("2020/11/24"), lty = 2) +
  annotate("text",label = as.Date("2020/11/24"),
           x = as.Date("2020/11/24"), y = .07, hjust = 1.1)+ 
  labs(x = "Date", y = "Cumulative deaths/cases ratio", 
       color = "Majority teaching posture",
       subtitle = "Yellow area represents the fall semester (08/18 - 12/15)") + 
  theme(legend.position = "bottom")+team_theme+scale_color_manual(values=col_theme)
```


## Pairwise
```{r}
ggplot(fall_cases,aes(y = death_incidence_per_1000, x = major_teaching)) + 
  geom_boxplot(aes(fill = major_teaching))+
  stat_compare_means(method = "anova")+ 
  stat_pvalue_manual(stat.test, label = "p",y.position = 2.5, step.increase = 0.15)+
  ylim(c(0,3.5))+ 
  theme_bw()+ 
  labs(y = "Death Incidence / 1,000 people",x = "",
       fill = "Majority teaching posture",
       title = "Death Incidence in the Fall Semester",
       caption = "Pairwise p-values come from Duncan pairwise comparison test") +
  theme(legend.position = "bottom",
        axis.text.x=element_blank())+team_theme+ scale_colour_manual(values=col_theme)+scale_fill_manual(values=col_theme)
#ggsave("boxplots.png", width = 7, height = 5)
```

# Appendix 3: Exponential growth model

## Data process

```{r}
cases_slope <- read.csv("county_splines.csv", header = T)%>%
  select(COUNTY,DATE,POPULATION,CUMDEATHS,log_tot_deaths,tot.slope,NEWDEATHS,rev_NEWDEATHS,log_new_deaths,new.slope)
# SHIFT THE DATE!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
cases_slope$DATE <- as.Date(cases_slope$DATE)-24
# get Majority teaching posture wide_teaching_enroll
cases_slope_teach <-death_teaching%>%
  select(-DATE,-POPULATION,-CUMDEATHS,-NEWDEATHS)%>%
  distinct()%>%
  right_join(cases_slope,by=c("COUNTY"))%>%
  filter(DATE>as.Date("2020-01-23"))
write.csv(cases_slope_teach,"cases_slope_teach.csv",row.names = F)
## ordering the teaching method factor to ensure the color order
cases_slope_teach$major_teaching <- factor(cases_slope_teach$major_teaching,levels = c("On Premises","Hybrid","Online Only"))
cases_slope_teach$DATE <- as.Date(cases_slope_teach$DATE)
```

## Select Max B1 & B0

```{r}
maxB1 <- cases_slope_teach%>%
  group_by(COUNTY)%>%
  filter(DATE >= as.Date("2020-08-18") & DATE<=as.Date("2020-12-15"))%>%
  summarise(max_B1 = max(new.slope), .groups = 'drop')

avgB1 <- cases_slope_teach%>%
  group_by(COUNTY)%>%
  filter(DATE >= as.Date("2020-08-18") & DATE<=as.Date("2020-12-15"))%>%
  summarise(avg_B1 = mean(new.slope), .groups = 'drop')
## avg3w_B0 ## average B0 of the first 3 weeks of school reopening 
## avg1w_2w_B0 ## OR average B0s between  2020-08-18 -7days and +14days [before the rate bounce back around the dashed line]
## avg3w_bf_B0 ## OR average B0s between  2020-08-18 -21days and 2020-08-18 [before the rate bounce back around the dashed line]
avgB0 <- cases_slope_teach%>%
  group_by(COUNTY)%>%
  filter(DATE > as.Date("2020-08-18") & DATE<as.Date(major_opendate)+21)%>%
  summarise(avg3w_B0 = mean(new.slope), .groups = 'drop')%>%
  left_join(cases_slope_teach%>%
  group_by(COUNTY)%>%
  filter(DATE > as.Date("2020-08-18")-7 & DATE<as.Date("2020-08-18")+14)%>%
  summarise(avg1w_2w_B0 = mean(new.slope)),by="COUNTY", .groups = 'drop')%>%
  left_join(cases_slope_teach%>%
  group_by(COUNTY)%>%
  filter(DATE < as.Date("2020-08-18") & DATE>=as.Date("2020-08-18")-21)%>%
  summarise(avg3w_bf_B0 = mean(new.slope)),by="COUNTY", .groups = 'drop')
#  B0 and B1
B0B1 <- death_teaching%>%
  distinct(COUNTY,POPULATION,NCHS.Urban.Rural.Status,Population.density)%>%
  left_join(maxB1,by="COUNTY")%>%
  left_join(wide_teaching_enroll, by = c("COUNTY" = "county"))%>%
  left_join(avgB1,by="COUNTY")%>%
  left_join(avgB0,by="COUNTY") %>%
  left_join(avg_mobility,by="COUNTY")
## ordering the teaching method factor to ensure the color order
B0B1$major_teaching <- factor(B0B1$major_teaching,levels = c("On Premises","Hybrid","Online Only"))
```


## Aggregate counties by teaching method

```{r}
cases_slope_teach_agg <- cases_slope_teach %>%
  drop_na(major_teaching)%>%
  group_by(DATE, major_teaching) %>%
  summarise(total_new_deaths = sum(rev_NEWDEATHS), .groups = "drop") %>%
  mutate(log_new_deaths = log(total_new_deaths + 1)) %>%
  group_by(major_teaching) %>%
  mutate(smooth.spline = smooth.spline(DATE,log_new_deaths,df = 398/28)$y,
         B = predict(smooth.spline(DATE,log_new_deaths,df = 398/28),deriv = 1)$y,B2 = predict(smooth.spline(DATE,log_new_deaths,df = 398/28),deriv = 2)$y)
week3_after_start <- as.Date("2020/08/18") + 21
ggplot(cases_slope_teach_agg, aes(x = DATE, color = major_teaching)) +
  geom_point(aes(y = log_new_deaths), alpha = .3)+ 
  geom_line(aes(y = smooth.spline), size = 1) +
  geom_rect(data = cases_slope_teach_agg[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "orange") + 
  labs(x = "Date", y = "Log ( New Deaths + 1 )", 
       color = "Majority teaching posture",
       caption = "Smoothing window set to every 4 weeks",
       subtitle = "Yellow area represents the fall semester (08/18 - 12/15)") + 
  geom_vline(xintercept = week3_after_start, lty = 2) + 
  annotate("text",label = "3 Weeks After",
           x = week3_after_start, y = 4, hjust = 1.1)+ 
  geom_vline(xintercept = as.Date("2020/08/18")+42, lty = 2) + 
  annotate("text",label = "6 Weeks After",
           x = as.Date("2020/08/18")+62, y = 4.7, hjust = 1.3)+ 
  theme(legend.position = "bottom")+team_theme+
  scale_colour_manual(values=col_theme)
####
ggplot(cases_slope_teach_agg, aes(x = DATE, color = major_teaching)) + 
  geom_line(aes(y = B), size = 1) +
  geom_rect(data = cases_slope_teach_agg[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "#E69F00") +
  geom_vline(xintercept = week3_after_start, lty = 2) + 
  annotate("text",label = "3 Weeks After",
           x = week3_after_start, y = .05, hjust = 1.1)+ 
  geom_vline(xintercept = as.Date("2020/08/18")+42, lty = 2) + 
  annotate("text",label = "6 Weeks After",
           x = as.Date("2020/08/18")+130, y = .06, hjust = 1.3)+ 
  labs(x = "Date", y = "Exponential Growth Coefficient", 
       color = "Majority teaching posture",
       caption = "Smoothing window set to every 4 weeks",
       subtitle = "Yellow area represents the fall semester (08/18 - 12/15)") + 
  theme(legend.position = "bottom")+team_theme+scale_color_manual(values=col_theme)
```


## Difference in B's

```{r}
B0w <- cases_slope_teach%>%
  filter(DATE==as.Date("2020-08-18"))%>%
  drop_na(major_teaching)%>%
  rename(new.slope0w=new.slope)
B3w <- cases_slope_teach%>%
  filter(DATE==as.Date("2020-08-18")+21)%>%
  drop_na(major_teaching)%>%
  rename(new.slope3w=new.slope)
B6w <- cases_slope_teach%>%
  filter(DATE==as.Date("2020-08-18")+42)%>%
  drop_na(major_teaching)%>%
  rename(new.slope6w=new.slope)
avg_mobi_0w3w <- case_mobility%>%
  left_join(major_reopening,by=c("COUNTY"))%>%
  group_by(COUNTY)%>%
  filter(DATE >= as.Date("2020-08-18")& DATE <as.Date("2020-08-18") + 21)%>%
  summarise(avg_full_work_prob = mean(full_work_prop_7d))
avg_mobi_3w6w <- case_mobility%>%
  left_join(major_reopening,by=c("COUNTY"))%>%
  group_by(COUNTY)%>%
  filter(DATE >= as.Date("2020-08-18")+ 21 & DATE <=as.Date("2020-08-18") + 42)%>%
  summarise(avg2_full_work_prob = mean(full_work_prop_7d))
B_diff <- B6w[,c(1:9,13,20)]%>%
  left_join(B3w%>%select(COUNTY,new.slope3w),by="COUNTY")%>%
  left_join(B0w%>%select(COUNTY,new.slope0w),by="COUNTY")%>%
  mutate(new.slope.diff = new.slope3w-new.slope0w,new.slope.diff2 = new.slope6w-new.slope3w)%>%
  left_join(avg_mobi_0w3w,by="COUNTY")%>%
  left_join(avg_mobi_3w6w,by="COUNTY")
B_diff$major_teaching <- factor(B_diff$major_teaching,levels = c("On Premises","Hybrid","Online Only"))
```

## max B vs. confounders

```{r}
na.omit(B0B1)%>%
  drop_na(major_teaching)%>%
  ggplot(aes(y=NCHS.Urban.Rural.Status,x=max_B1,fill=major_teaching))+
  facet_grid(~major_teaching)+
  geom_boxplot()+
  team_theme+theme(legend.position = "bottom")+
  labs(x="Max B",y="Urban-Rural Status",title="Distribution of Max B",
       fill="Majority teaching posture")+
  scale_fill_manual(values=col_theme)+
  team_theme

ggsave("maxB_urban.png",width = 10.5,height = 5)


##### Micropolitan counties

maxB_major_teaching.aov <- aov(max_B1 ~ major_teaching,data = na.omit(B0B1)%>%
  filter(NCHS.Urban.Rural.Status=="Micropolitan"))
summary(maxB_major_teaching.aov)

# p-value of .012
stat.test <- PostHocTest(maxB_major_teaching.aov, method = "duncan")$major_teaching %>%
  as.data.frame()%>%
  rownames_to_column("group") %>%
  separate(group,"-", into = c("group1","group2")) %>%
  mutate(pval = round(pval,3),
         p = case_when(pval <= .01~ "**",
                       pval <= .05 ~ "*",
                       TRUE ~ "NS"))%>%
  select(group1, group2, pval, p)



na.omit(B0B1)%>%
  filter(NCHS.Urban.Rural.Status=="Micropolitan")%>%
  drop_na(major_teaching)%>%
  ggplot(aes(x=major_teaching,y=max_B1))+
  geom_boxplot(aes(fill=major_teaching))+
  #ylim(c(0,0.05))+
  #stat_compare_means(method = "anova",size=6,label.y.npc=0.96,label.x.npc = 0.4)+ 
  #stat_pvalue_manual(stat.test, label = "p",y.position = 0.03, step.increase = 0.15,
   #                  size = 6,bracket.nudge.y = 0.001)+
  team_theme+
  theme(legend.position = " ")+
  labs(y="Max Growth Coefficient B",x="Majority teaching posture",title="Distribution of Maximum Growth Coefficient \nin Micropolitan Counties",
       fill="Majority teaching posture")+
  scale_fill_manual(values=col_theme)

ggsave("maxB_micro.png",width = 6.5,height = 6)



na.omit(B0B1)%>%
  drop_na(major_teaching)%>%
  ggplot(aes(x=major_teaching,y=max_B1))+
  geom_boxplot(aes(fill=major_teaching))+
  #ylim(c(0,0.05))+
  #stat_compare_means(method = "anova",size=6,label.y.npc=0.96,label.x.npc = 0.4)+ 
  #stat_pvalue_manual(stat.test, label = "p",y.position = 0.03, step.increase = 0.15,
   #                  size = 6,bracket.nudge.y = 0.001)+
  team_theme+
  theme(legend.position = " ")+
  labs(y="Max Growth Coefficient B",x="Majority teaching posture",title="Distribution of Maximum Growth Coefficient",
       fill="Majority teaching posture")+
  scale_fill_manual(values=col_theme)

ggsave("maxB.png",width = 6.5,height = 6)
```

## Max B vs. Mobility and Population Density
```{r}

na.omit(B0B1)%>%
  drop_na(major_teaching)%>%
  ggplot(aes(x=avg_full_work_prob,y=max_B1,group=major_teaching,color=major_teaching))+
  geom_point(aes(size=log(Population.density),alpha=log(Population.density)))+
  geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+
  labs(y="Max B",x="Averaged \n%6hr+ Away Home in Fall semester",title="Max B in Fall v.s. Average Mobility \nAll Counties",color="Majority teaching posture",size = "Log of Population Density",alpha= "Log of Population Density" )+
  team_theme+theme(legend.position = "bottom")+
  scale_color_manual(values=col_theme)
  
ggsave("maxB_mob.png",width = 6.5,height = 6)
######

na.omit(B0B1)%>%
  drop_na(major_teaching)%>%
  ggplot(aes(x=log(Population.density),y=max_B1,group=major_teaching,color=major_teaching))+
  geom_point(aes(size=avg_full_work_prob,alpha=avg_full_work_prob))+
  geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+
  labs(y="Max B",x="Log of Population Density",title="Max B in Fall v.s. Log of Population Density \nAll Counties",color="Majority teaching posture",size = "Averaged \n%6hr+ Away Home in Fall semester",alpha= "Averaged \n%6hr+ Away Home in Fall semester" )+team_theme+theme(legend.position = "bottom")+
  scale_color_manual(values=col_theme)


ggsave("maxB_pop.png",width = 6.5,height = 6)
```

## Max B vs. Mobility and Population Density in Micropolitan counties
```{r}

na.omit(B0B1)%>%
  filter(NCHS.Urban.Rural.Status=="Micropolitan")%>%
  drop_na(major_teaching)%>%
  ggplot(aes(x=avg_full_work_prob,y=max_B1,group=major_teaching,color=major_teaching))+
  geom_point(aes(size=log(Population.density),alpha=log(Population.density)))+
  #geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+
  labs(y="Max B",x="Averaged \n%6hr+ Away Home in Fall semester",
       title="Max B in Fall v.s. Average Mobility \nOnly Micropolitan Counties",
       color="Majority teaching posture",
       size = "Log of Population Density",
       alpha= "Log of Population Density" )+
  team_theme+theme(legend.position = "bottom")+
  scale_color_manual(values=col_theme)+  
  guides(
    size = guide_legend(order = 1),
    alpha = guide_legend(order = 1),
    fill = guide_legend(order = 0)
  )
  
ggsave("maxB_mob_micro.png",width = 6.5,height = 6)
######

na.omit(B0B1)%>%
  filter(NCHS.Urban.Rural.Status=="Micropolitan")%>%
  drop_na(major_teaching)%>%
  ggplot(aes(x=log(Population.density),y=max_B1,group=major_teaching,color=major_teaching))+
  geom_point(aes(size=avg_full_work_prob,alpha=avg_full_work_prob))+
  #geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+
  labs(y="Max B",x="Log of Population Density",
       title="Max B in Fall v.s. Log of Population Density \nOnly Micropolitan Counties",
       color="Majority teaching posture",
       size = "Averaged \n%6hr+ Away Home in Fall semester",
       alpha= "Averaged \n%6hr+ Away Home in Fall semester" )+
  team_theme+theme(legend.position = "bottom")+
  scale_color_manual(values=col_theme)


ggsave("maxB_pop_micro.png",width = 6.5,height = 6)
```

 
## B6B3 vs. B3B0

```{r}
B_diff%>%
  ggplot(aes(x=new.slope.diff,y=new.slope.diff2,group=major_teaching,color=major_teaching))+geom_point(aes(size=log(Population.density),alpha=log(Population.density)))+geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+team_theme+
  labs(y="B(6w) - B(3w)",x="B(3w) - B(0w)",
       title="Change in growth before vs. after school reopen\nAll Counties",
       color="Majority teaching posture",
       size = "Log of Population Density",
       alpha= "Log of Population Density" ,fill="Majority teaching posture")+
  scale_color_manual(values=col_theme)+theme(legend.position = "bottom")+
  geom_hline(yintercept = 0,lty=2)+geom_vline(xintercept = 0,lty=2)

ggsave("B63vsB30.png",width = 6.5,height = 6.5)

B_diff_micro%>%
  ggplot(aes(x=new.slope.diff,y=new.slope.diff2,group=major_teaching,color=major_teaching))+geom_point(aes(size=log(Population.density),alpha=log(Population.density)))+
  geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+team_theme+
  labs(y="B(6w) - B(3w)",x="B(3w) - B(0w)",
       title="Change in growth before vs. after school reopen\nOnly Micropolitan Counties",
       color="Majority teaching posture",
       size = "Log of Population Density",
       alpha= "Log of Population Density" ,fill="Majority teaching posture")+
  scale_color_manual(values=col_theme)+theme(legend.position = "bottom")+
  geom_hline(yintercept = 0,lty=2)+geom_vline(xintercept = 0,lty=2)

ggsave("B63vsB30_micro.png",width = 6.5,height = 6.5)
```


## Acceleration vs. density

```{r}
# pop density plot
B_diff%>%
  drop_na(major_teaching)%>%
  ggplot(aes(x=log(Population.density),y=new.slope.diff2,group=major_teaching,color=major_teaching))+geom_point(aes(size=avg2_full_work_prob,alpha=avg2_full_work_prob))+geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+theme_minimal()+team_theme+team_theme+labs(y="B(6w) - B(3w)",x="Log of Population",title="Growth in 3w~6w vs Log of Population Density",color="Majority teaching posture",size = "3w~6w\nAveraged %6hr+ Away Home",alpha= "3w~6w\nAveraged %6hr+ Away Home" )+scale_color_manual(values=col_theme)+theme(legend.position = "bottom")
# mobility plot
B_diff%>%
  drop_na(major_teaching)%>%
  ggplot(aes(x=avg2_full_work_prob,y=new.slope.diff2,group=major_teaching,color=major_teaching))+geom_point(aes(size=log(Population.density),alpha=log(Population.density)))+geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+theme_minimal()+team_theme+labs(y="B(6w) - B(3w)",x="Average mobility proportion during fall semester",title="Growth in 3w~6w vs Average full mobility proportion in 3w~6w\n Averaged %6hr+ Away Home",size="Log of Population Density",alpha="Log of Population Density",color="Majority teaching posture")+scale_color_manual(values=col_theme)+theme(legend.position = "bottom")
```

## Micropolitan Counties

```{r}
# Pop density vs RURAL
ohio_profile%>%
  left_join(wide_teaching_enroll[,c("county","major_teaching")],
            by = c("County" = "county"))%>%
  drop_na(major_teaching)%>%
  ggplot(aes(y=NCHS.Urban.Rural.Status,x=log(Population.density),
             fill=major_teaching))+
  facet_grid(~major_teaching)+
  geom_boxplot()+
  labs(fill="Majority Teaching \nMethod",
       size="Averaged \n%6hr+ Away Home",
       title="Distribution of Log Population Density \nby Rural-urban Status",
       x="Log Population Density", y= "Rural-urban Status")+
  team_theme+
  scale_fill_manual(values=col_theme)+
  theme(legend.position = "bottom")
ggsave("pop_dist.png", width = 10, height = 5)

# diff B1 vs RURAL
B_diff%>%
  drop_na(major_teaching)%>%
  ggplot(aes(y=NCHS.Urban.Rural.Status,x=new.slope.diff2,fill=major_teaching))+
  facet_grid(~major_teaching)+
  geom_boxplot()+
  labs(fill="Majority teaching posture",size="Averaged \n%6hr+ Away Home",
       title="Distribution of B(6 weeks after) minus B(3 weeks after) \nby Rural-urban Status",
       x="B(6w) - B(3w)", y= "Rural-urban Status")+
  team_theme+scale_fill_manual(values=col_theme)+theme(legend.position = "bottom")
```


### information about 7 mircopolitan counties on-premises
```{r}
B_diff_micro%>%
  filter(major_teaching=="On Premises")
```


```{r}
# Micropolitan map plots
ohio_profile%>%
  distinct(County,NCHS.Urban.Rural.Status) %>%
  mutate(is_micro = factor(ifelse(NCHS.Urban.Rural.Status == "Micropolitan",1,0)))%>%
  left_join(ohio_map,by=c('County'='county'))%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = is_micro),
               color = "gray") + 
  coord_fixed(1.3) + theme_map() +
  scale_fill_manual(values = c("0" ="#0072B2",
                    "1" ="#009E73"))+
  labs(fill='Micropolitan Counties')+
  theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "bottom",
        legend.key.size = unit(2,"lines"))

# Micropolitan map plots
ohio_profile%>%
  distinct(County,NCHS.Urban.Rural.Status) %>%
  left_join(wide_teaching_enroll[,c("county","major_teaching")], 
            by = c("County" = "county"))%>%
  mutate(is_micro = factor(ifelse(NCHS.Urban.Rural.Status == "Micropolitan",1,0)),
         micro_teach = factor(ifelse(is_micro == 1, major_teaching, "Not Micro")))%>%
  left_join(ohio_map,by=c('County'='county'))%>%
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = micro_teach),
               color = "gray",alpha=0.9) + 
  coord_fixed(1.3) + theme_map() +
  scale_fill_manual(values = c(col_theme, "Not Micro" = "white"))+
  labs(fill='Micropolitan \nCounties')+
  theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "bottom",
        legend.key.size = unit(2,"lines"))
#ggsave("micro.png", width = 7, height = 5)
```


```{r}
## Filter only the micro counties
B_diff_micro <- B_diff%>%
  drop_na(major_teaching)%>%
  filter(NCHS.Urban.Rural.Status=="Micropolitan") %>%
  mutate(diff_in_slope = new.slope.diff2 - new.slope.diff)

#selected county for label
onpremise_text <- B_diff_micro%>%
  filter(major_teaching=="On Premises")%>%
  filter(COUNTY %in% c("CHAMPAIGN","DEFIANCE","MERCER","VAN WERT"))%>%
  mutate(County = str_to_title(COUNTY))
```


## Acceleration vs. mob 

```{r}
B_diff_micro%>%
  ggplot(aes(x=avg_full_work_prob,y=new.slope.diff,group=major_teaching,color=major_teaching))+geom_point(aes(size=log(Population.density),alpha=log(Population.density)))+geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+
  labs(y="Change in Growth",x="Average mobility proportion in 0w-3w",
       title="Difference in slope of B(0w) and B(3w)\nOnly Micropolitan Counties",
       color="Majority teaching posture",
       size = "0w~3w\nAveraged %6hr+ Away Home",
       alpha= "0w~3w\nAveraged %6hr+ Away Home")+
  scale_color_manual(values=col_theme)+
  theme(legend.position = "bottom")+
  team_theme+
  geom_text(data=onpremise_text,aes(label=County),
            color='black',size=4,hjust=0.6, vjust=-0.1)
ggsave("acc_mob_micro.png",width = 6.5,height = 6)


B_diff_micro%>%
  ggplot(aes(x=avg2_full_work_prob,y=new.slope.diff2,group=major_teaching,color=major_teaching))+geom_point(aes(size=log(Population.density),alpha=log(Population.density)))+geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+
  labs(y="Change in Growth",x="Average mobility proportion in 0w-3w",
       title="Difference in slope of B(3w) and B(6w)\nOnly Micropolitan Counties",
       color="Majority teaching posture",
       size = "0w~3w\nAveraged %6hr+ Away Home",
       alpha= "0w~3w\nAveraged %6hr+ Away Home")+
  scale_color_manual(values=col_theme)+theme(legend.position = "bottom")+
  team_theme+
  geom_text(data=onpremise_text,aes(label=County),
            color='black',size=4,hjust=0.6, vjust=-0.1)
ggsave("acc2_mob_micro.png",width = 6.5,height = 6)
```

## Acceleration vs. pop

```{r}
B_diff_micro%>%
  ggplot(aes(x=log(Population.density),y=new.slope.diff,
             group=major_teaching,color=major_teaching))+
  geom_point(aes(size=avg_full_work_prob,alpha=avg_full_work_prob))+
  geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+
  team_theme+
  labs(y="Acceleration",x="Log of Population Density",
       title="Difference in slope of B(0w) and B(3w)\nOnly Micropolitan Counties",
       color="Majority teaching posture",
       size = "0w~3w\nAveraged %6hr+ Away Home",
       alpha= "0w~3w\nAveraged %6hr+ Away Home" ,fill="Majority teaching posture")+
  scale_color_manual(values=col_theme)+
  theme(legend.position = "bottom")+
  geom_text(data = onpremise_text,aes(label=County),
            color='black',size=4,hjust=0.6, vjust=-0.8)

ggsave("acc_pop_micro.png",width = 7.5,height = 6)


B_diff_micro%>%
  ggplot(aes(x=log(Population.density),y=new.slope.diff2,
             group=major_teaching,color=major_teaching))+
  geom_point(aes(size=avg2_full_work_prob,alpha=avg2_full_work_prob))+
  geom_smooth(method = "lm", se=F, formula = y ~ x,alpha=0.1)+
  team_theme+
  labs(y="Acceleration",x="Log of Population Density",
       title="Difference in slope of B(3w) and B(6w)\nOnly Micropolitan Counties",
       color="Majority teaching posture",
       size = "3w~6w\nAveraged %6hr+ Away Home",
       alpha= "3w~6w\nAveraged %6hr+ Away Home" ,fill="Majority teaching posture")+
  scale_color_manual(values=col_theme)+
  theme(legend.position = "bottom")+
  geom_text(data=onpremise_text,aes(label=County),color='black',
            size=4,hjust=0.6, vjust=-0.1)

ggsave("acc2_pop_micro.png",width = 7.5,height = 6)

```



```{r}
B_diff_micro%>%
  ggplot(aes(x=log(Population.density),y=diff_in_slope,
             group=major_teaching,color=major_teaching))+
  geom_point(aes(size=avg2_full_work_prob,alpha=avg2_full_work_prob))+
  geom_smooth(method = "lm", se=F,
              formula = y ~ x,alpha=0.1)+
  team_theme+
  labs(y="Acceleration",x="Log of Population Density",
       title="Difference in slope of B(0w) and B(3w)\nOnly Micropolitan Counties",
       size = "3w~6w\nAveraged %6hr+ Away Home",
       alpha= "3w~6w\nAveraged %6hr+ Away Home",
       color="Majority teaching posture")+
  scale_color_manual(values=col_theme)+theme(legend.position = "bottom")
#ggsave("acc.png", width = 7, height = 5)
```

## 3 colored boxplots for response variable (total number of deaths before semester)/pop

```{r}
death_semester <- cases%>%
  filter(DATE>="2020-08-18" & DATE<="2020-12-15")%>%
  group_by(COUNTY)%>%
  filter(row_number()==1 | row_number()==n())%>%
  summarize(death_semester = diff(CUMDEATHS),DATE = DATE,cdeaths=CUMDEATHS,label=c("start","end"))%>%
  dcast(COUNTY+death_semester~label,value.var = "cdeaths")

death_semester%>%
  left_join(B0B1%>%select(COUNTY,major_teaching),by="COUNTY")%>%
  left_join(death_prop%>%select(COUNTY,POPULATION),by="COUNTY")%>%
  mutate(d_inci_before_sem=start/POPULATION*1000)%>%
  na.omit()%>%
  ggplot(aes(x=major_teaching,y=d_inci_before_sem,group=major_teaching,fill=major_teaching))+geom_boxplot()+theme_minimal()+team_theme+
  labs(y="Death Incidence",x="Teaching Method",title="Death Incidence vs Teaching Method \n-(total number of deaths before semester)/pop*1000",color="Teaching Method")+team_theme+scale_fill_manual(values=col_theme)+scale_color_manual(values=col_theme)+theme(legend.position = "bottom")
ggsave("boxplotmaxB1vstmethod_bfs.png",width = 7, height = 5)
```


```{r}
ratio <- cases%>%
  left_join(wide_teaching_enroll,by=c("COUNTY"="county"))%>%
  drop_na(major_teaching)%>%
  group_by(major_teaching,DATE)%>%
  summarise(ratios = sum(CUMDEATHS)/sum(CUMCONFIRMED))
ratio%>%
  ggplot(aes(x=DATE,y=ratios,color=major_teaching))+geom_line(size=1)+
  geom_rect(data = ratio[1,],
            aes(xmin=as.Date("2020/08/18"), xmax=as.Date("2020/12/15"),
                ymin=-Inf,ymax=Inf),
            color = NA,alpha=0.2, show.legend = F, fill = "orange") + 
  geom_vline(xintercept = as.Date("2020/11/24"), lty = 2) +
  annotate("text",label = as.Date("2020/11/24"),
           x = as.Date("2020/11/24"), y = .07, hjust = 1.1)+ 
  labs(x = "Date", y = "Cumulative deaths/cases ratio", 
       color = "Majority teaching posture",
       subtitle = "Yellow area represents the fall semester (08/18 - 12/15)") + 
  theme(legend.position = "bottom")+team_theme+scale_color_manual(values=col_theme)
```

```{r}
teaching_profile%>%
  ggplot(aes(x=major_teaching,y=Percent.uninsured))+geom_boxplot(aes(fill=major_teaching),width=0.6)+
  stat_compare_means(method = "anova",size=6,label.y.npc=0.85)+ 
  stat_pvalue_manual(stat.test, label = "p",y.position = 1, step.increase = 0.15,size = 6,bracket.nudge.y = 8)+
  labs(title="Percent of Uninsured Population by Teaching Method",x="Majority teaching posture",y="Percent of uninsured population",caption = "Post-hoc pairwise testing with Duncan Method")+
  team_theme+theme(legend.position = "")+scale_fill_manual(values=col_theme)

ggsave("p2.png",width = 6.5,height = 6)

senior_major_teaching.aov <- aov(Percent.Population.65..yrs~ major_teaching,data = teaching_profile)

summary(senior_major_teaching.aov)

# p-value of .012
stat.test <- PostHocTest(senior_major_teaching.aov, method = "duncan")$major_teaching %>%
  as.data.frame()%>%
  rownames_to_column("group") %>%
  separate(group,"-", into = c("group1","group2")) %>%
  mutate(pval = round(pval,3),
         p = case_when(pval <= .01~ "**",
                       pval <= .05 ~ "*",
                       TRUE ~ "NS"))%>%
  select(group1, group2, pval, p)

teaching_profile%>%
  ggplot(aes(x=major_teaching,y=Percent.Population.65..yrs))+
  geom_boxplot(aes(fill=major_teaching),width=0.6)+
  stat_compare_means(method = "anova",size=6,label.y.npc=0.95)+ 
  stat_pvalue_manual(stat.test, label = "p",y.position = 1, step.increase = 0.15,size = 5,bracket.nudge.y = 2)+
  labs(title="Percent of 65years+ Population by Teaching Method",x="Major Teaching Method",y="Percent of 65+yrs population",fill="Majority teaching posture",caption = "Post-hoc pairwise testing with Duncan Method")+
  team_theme+theme(legend.position = "")+
  scale_fill_manual(values=col_theme)

```


# Appendix 4.

## Gamma distribution for the time lengths from infections to deaths

We know from previous study that the mean for this Gamma distribution is 23.9, with a coefficient of variation being 0.4.


```{r}
# package for The Gamma Distribution (Alternative Parameterization)
# install.packages("EnvStats")
library(EnvStats)

time_to_deaths <- 1:50
prob_time_to_deaths <- dgammaAlt(x = time_to_deaths,mean = 23.9, cv = 0.4)
## shift x 
gamma_plot <- data.frame(prob_time_to_deaths,time_to_deaths,time_to_deaths+5,time_to_deaths+10,time_to_deaths+15)
colnames(gamma_plot) <- c("prob","time1","time2","time3","time4")

ggplot(gamma_plot)+
  geom_line(aes(x=time1,y=prob),colour = "black")+
  geom_vline(xintercept = 5.2,lty=2,colour="darkgreen")+
  geom_vline(xintercept = 15.2,lty=2,colour="darkgreen")+
  geom_vline(xintercept = 25.2,lty=2,colour="darkgreen")+
  labs(x="Time from infections to deaths",
       y="Probability of died after x days")+team_theme+theme(legend.position = "bottom")

ggsave("gamma.png",width = 6,height = 4)
```


